### 2016-07-15 presentation.tex
starting presentation, comments through pg 12

\documentclass[pdf]{beamer}
\mode<presentation>{}

\title{Evaluating the Capabilities of Lattice Boltzmann Method for non-Newtonian and Free-surface Flows towards Applications in Wellbore Cementing}
\author{Matthew Grasinger}
\date{July 21st, 2016}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\newcommand{\pw1}{\linewidth/3}

\begin{frame}{non-Newtonian Flow}
  \begin{tabular}{cc}
    \includegraphics[width=\pw1]{figs/slurry-backfill} & \includegraphics[width=\pw1]{figs/molten-metal} \\
    \includegraphics[width=\pw1]{figs/varnish} & \includegraphics[width=\pw1]{figs/blood-circulation}
\end{tabular}
\end{frame}

\end{document}