### 2016-07-19 src/algorithm/streaming.tex
writing pseudo for algorithms

\begin{algorithm}
\caption{Streaming} \label{algo:streaming}
\begin{algorithmic}[1]
\Procedure{streaming}{$f$}
Initialize temporary 3D array the size of lattice, $f_temp$
\EndProcedure
\end{algorithmic}
\end{algorithm}