### 2016-07-27 src/algorithm/update-cell-states.tex

\begin{algorithm}
\caption{Update Cell States and Mass Redistribution} \label{algo:update-cell-states}
\begin{algorithmic}[1]
\Procedure{UpdateStates}{}

\Statex
\State Initialize (empty) data structure, $\iset_f$ \Comment{$\iset$ cells to be transitioned to $\fset$}
\State Initialize (empty) data structure, $\iset_g$ \Comment{$\iset$ cells to be transitioned to $\gset$}
\State Initialize variable that represents global excess mass, $M_{gex} = 0$

\Statex
\ForAll{cells $\pos \in \iset$} \Comment{Mark interface cells to transition}
  \If{$M(\pos, t) < -\delta_{trans}$}
    \State Add $\pos$ to $\iset_f$
  \ElsIf{$M(\pos, t) > \rho(\pos, t) + \delta_{trans}$}
    \State Add $\pos$ to $\iset_g$
  \EndIf
\EndFor

\Statex
\ForAll{cells $\pos \in \iset_f$}
\Statex \Comment{Prepare neighborhoods of $\iset$ cells that are to be transitioned to $\fset$ cells}
  \State Calculate the average density, $\rho_{avg}$, of the neighborhood
  \State Calculate the average velocity, $\mathbf{u}_{avg}$, of the neighborhood
  \ForAll{cells $\pos_n$ in the neighborhood of $\pos$}
    \If{$\pos_n \in \gset$}
      \State Transition $\pos_n$ from a $\gset$ state to $\iset$ state
      \State $f_i(\pos_n, t) = f_i^{eq}(\rho_{avg}, \mathbf{u}_{avg})$
    \ElsIf{$\pos_n \in \iset_g$}
      \State Remove $\pos_n$ from $\iset_g$ \Comment{Ensure interface is continuous}
    \EndIf
  \EndFor
\EndFor

\Statex
\ForAll{cells $\pos \in \iset_g$}
\Statex \Comment{Prepare neighborhoods of $\iset$ cells that are to be transitioned to $\gset$ cells}
  \ForAll{cells $\pos_n$ in the neighborhood of $\pos$}
    \If{$\pos_n \in \fset$}
      \State Transition $\pos_n$ from a $\fset$ state to $\iset$ state
    \EndIf
  \EndFor
\EndFor

\algstore{ucsalg1}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\begin{algorithmic}[1]
\algrestore{ucsalg1}
\ForAll{cells $\pos \in \iset_f$} 
\Statex \Comment{Transition from $\iset$ to $\fset$ and redistribute excess mass}
  \State Let $M_{ex} = M(\pos, t) - \rho(\pos, t)$
  \State Approximate the unit normal from the interface toward empty space, \indent \hspace{0.3in} $\mathbf{n} = \frac{1}{2} \begin{Bmatrix}
\epsilon_{i-1, j} - \epsilon_{i+1, j} \\
\epsilon_{i, j-1} - \epsilon_{i, j+1}
\end{Bmatrix}$ \Comment{Central difference of fluid fraction}
  \State Initialize $v_{sum} = 0$
  \State Initialize (empty) data structure, $\iset_r$
  \ForAll{cells $\pos_n$ in the neighborhood of $\pos$}
    \If{$\pos_n \in \iset$}
      \State Add $\pos_n$ to $\iset_r$
      \State Let $v_i = \begin{cases}
					\mathbf{n} \cdot \pvel_i, & \mathbf{n} \cdot \pvel_i > 0 \\
					0, & \text{otherwise}
				\end{cases}$
      \State $v_{sum} = v_{sum} + v_i$
    \EndIf
  \EndFor
  \If{$v_{sum} > 0$}
    \ForAll{cells $\pos_j \in \iset_r$}
      \State $M(\pos_j, t) = M(\pos_j, t) + M_{ex} \frac{v_i}{v_{sum}}$
    \EndFor
  \ElsIf{$v_{sum} = 0$ and $\iset_r \neq \emptyset$}
    \State Let $n_r$ be the number of cells in $\iset_r$
    \ForAll{cells $\pos_j \in \iset_r$}
      \State $M(\pos_j, t) = M(\pos_j, t) + \frac{M_{ex}}{n_r}$
    \EndFor
  \Else
    \State $M_{gex} = M_{gex} + M_{ex}$
  \EndIf
  \State Transition $\pos$ from an $\iset$ state to $\fset$ state
  \State Set $M(\pos, t) = \rho(\pos, t)$
\EndFor

\algstore{ucsalg2}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\begin{algorithmic}[1]
\algrestore{ucsalg2}
\ForAll{cells $\pos \in \iset_g$}
\Statex \Comment{Transition from $\iset$ to $\gset$ and redistribute excess mass}
  \State Let $M_{ex} = M(\pos, t)$
  \State Approximate the unit normal from the interface toward empty space, \indent \hspace{0.3in} $\mathbf{n} = \frac{1}{2} \begin{Bmatrix}
\epsilon_{i-1, j} - \epsilon_{i+1, j} \\
\epsilon_{i, j-1} - \epsilon_{i, j+1}
\end{Bmatrix}$ \Comment{Central difference of fluid fraction}
  \State Initialize $v_{sum} = 0$
  \State Initialize (empty) data structure, $\iset_r$
  \ForAll{cells $\pos_n$ in the neighborhood of $\pos$}
    \If{$\pos_n \in \iset$}
      \State Add $\pos_n$ to $\iset_r$
      \State Let $v_i = \begin{cases}
					-\mathbf{n} \cdot \pvel_i, & \mathbf{n} \cdot \pvel_i < 0 \\
					0, & \text{otherwise}
				\end{cases}$
      \State $v_{sum} = v_{sum} + v_i$
    \EndIf
  \EndFor
  \If{$v_{sum} > 0$}
    \ForAll{cells $\pos_j \in \iset_r$}
      \State $M(\pos_j, t) = M(\pos_j, t) + M_{ex} \frac{v_i}{v_{sum}}$
    \EndFor
  \ElsIf{$v_{sum} = 0$ and $\iset_r \neq \emptyset$}
    \State Let $n_r$ be the number of cells in $\iset_r$
    \ForAll{cells $\pos_j \in \iset_r$}
      \State $M(\pos_j, t) = M(\pos_j, t) + \frac{M_{ex}}{n_r}$
    \EndFor
  \Else
    \State $M_{gex} = M_{gex} + M_{ex}$
  \EndIf
  \State Transition $\pos$ from an $\iset$ state to $\gset$ state
  \State Set $M(\pos, t) = 0$
\EndFor

\Statex
\State Let $n_i$ be the number of cells in $\iset$
\ForAll{cells $\pos \in \iset$}
\Statex \Comment{Redistribute remaining excess mass evenly among interface cells}
  \State $M(\pos, t) = M(\pos, t) + \frac{M_{gex}}{n_i}$
\EndFor

\EndProcedure
\end{algorithmic}
\end{algorithm}