### 2016-08-11 thesis.tex

% MG: replace \~\cite with ~\cite to ensure citations stay on same line
% MG: need an overall consistency in 'flucuation' vs. 'oscillation' vs. 'noise'; use same word if you mean same thing.
% MG: same note as above for 'local equilibrium' vs. 'quasiequilibrium'
% MG: consider using collision frequency, \omega, instead of relaxation time, \tau, so that the relaxation time is not confused with shear stress

\documentclass[pdftex,ms]{pittetd}

\usepackage[pdftex]{graphicx}

\usepackage[affil-it]{authblk}
\usepackage{lineno}
\usepackage{soul}
\usepackage{color}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{enumitem}
\setlistdepth{9}

\newlist{myEnumerate}{enumerate}{9}
\setlist[myEnumerate,1]{label=(\arabic*)}
\setlist[myEnumerate,2]{label=(\Roman*)}
\setlist[myEnumerate,3]{label=(\Alph*)}
\setlist[myEnumerate,4]{label=(\roman*)}
\setlist[myEnumerate,5]{label=(\alph*)}
\setlist[myEnumerate,6]{label=(\arabic*)}
\setlist[myEnumerate,7]{label=(\Roman*)}
\setlist[myEnumerate,8]{label=(\Alph*)}
\setlist[myEnumerate,9]{label=(\roman*)}

%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bm}
\usepackage[plain]{fancyref}
\usepackage{tabulary}
\usepackage{pbox}
\usepackage{algorithm}
\usepackage{algpseudocode}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\def\mkfancyprefix#1#2{%
  \@namedef{fancyref#1labelprefix}{#1}%
  \begingroup\def\x{\endgroup\frefformat{plain}}%
    \expandafter\x\csname fancyref#1labelprefix\endcsname
      {\MakeLowercase{#2}\fancyrefdefaultspacing##1}%
  \begingroup\def\x{\endgroup\Frefformat{plain}}%
    \expandafter\x\csname fancyref#1labelprefix\endcsname
      {#2\fancyrefdefaultspacing##1}%
}
\makeatother
\mkfancyprefix{algo}{Algorithm}

\patch{amsmatch}
\patch{amsthm}

\input{src/notation.tex}

\title{Evaluating the Capabilities of the Lattice Boltzmann Method for non-Newtonian and Free-surface Flow towards Applications in Wellbore Cementing}
% The optional argument is the version of the title that will appear in Acrobat Reader's Document Info dialog box.
\author{Matthew Grasinger}
\degree{BS in Civil Engineering, University of Pittsburgh, Pittsburgh, PA 2013}
\date{July 21, 2016}
\keywords{lattice Boltzmann method, non-Newtonian flow, free surface flow, primary cementing}
\subject{Final}
\school{Swanson School of Engineering}
\degreesought{Master of Science}

\chapterfloats

\begin{document}
\maketitle

% For the committee membership page, you have to provide the names and affiliations of the members. The first one will
% be treated by pittetd as the committee chair (thesis/dissertation advisor).
\committeemember{John C. Brigham, PhD, Associate Professor, Department of Civil and Environmental Engineering}
\committeemember{Julie M. Vandenbossche, PhD, Associate Professor, Department of Civil and Environmental Engineering}
\committeemember{Anthony T. Iannacchione, PhD, Associate Professor, Department of Civil and Environmental Engineering}

\school{Swanson School of Engineering}
\makecommittee
\copyrightpage

\begin{abstract}

When oil and gas wellbores are drilled, barriers must be put in place to ensure that fluids do not leak out of the wellbore.
Wellbore leakage can lead to environmental damage, loss of pressure at the wellhead, and consequently, loss of production.
An important yet vulnerable barrier is the cement annulus.
Every well has unique subsurface conditions, and so no cement slurry mix design both performs well and is economical for cementing all wells.
The lattice Boltzmann method (LBM) is a promising technique for simulating primary wellbore cementing because it is well-suited for efficiently simulating non-Newtonian flows, multiphase multicomponent flows, and flows through complex geometries--namely, some of the complexities associated with the mechanics of primary cementing.

Despite the advantages of LBM, there are considerations that must be made, as with all computational methods, in regards to the accuracy and numerical stability of the solution.
Issues with accuracy and numerical stability are especially relevant in the flow of non-Newtonian fluids because of the nonlinear constitutive relationship between shear stress and strain-rate.
\Fref{chap:lbm-stability} is a numerical investigation of the accuracy, stability, and computational efficiency of different LB methods in simulating non-Newtonian fluid flow.
The accuracy and computational time are presented in \Fref{sec:numerical-study} for various LB methods applied to two different benchmark problems, Poiseuille flow and lid-driven cavity flow, with two different non-Newtonian constitutive models, the Bingham plastic constitutive relationship and a power-law constitutive relationship.

Once the ground-work has been laid for simulating non-Newtonian flows using LBM, the algorithm was extended in \Fref{chap:lbm-cement} to incorporate simulation of free-surface flow.
The extended LBM model was used to study primary cementing of a dry annulus, i.e. an annulus that is not filled with drilling mud.
More specifically, the study involved defining different cement slurry flows and investigating how well each slurry flow filled different wellbore geometries.
The study is followed by conclusions and a discussion of future work.

\end{abstract}

\tableofcontents
\listoftables
\listoffigures
%\listofalgorithms

\preface
I would like to thank my wife and best friend, Shelby, my parents, and my family.
Their love, support, and example means everything to me.
I would like to thank the best advisor ever, Dr. John Brigham, for his patience and for all that he has taught me. 
I would also like to thank the other members of my committee, Dr. Julie Vandenbossche, and Dr. Anthony Iannacchione, both of whom have been great guides and teachers to me.
Lastly, I would like to thank my fellow students, especially Rob, Jing, Mo, and Scott, for helping me to learn along the way and for the comradery that every graduate student needs.

\chapter{Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Models in Simulating Non-Newtonian Fluid Flow} \label{chap:lbm-stability}

\input{src/stability-paper.tex}

\chapter{Free Surface and non-Newtonian Simulation of Wellbore Cementing using the Lattice Boltzmann Method} \label{chap:lbm-cement}

\input{src/cementing-paper.tex}

\chapter{Conclusions}

Various Lattice Boltzmann methods were investigated for their efficacy in simulating primary cementing.
Although non-Newtonian fluid mechanics only represent a portion of the challenges associated with simulating primary cementing, the fluid components that can be present in primary cementing exhibit various non-Newtonian behaviors such as yield stress or Bingham plastic behavior (e.g. cement slurry), power-law behavior (e.g. drilling mud), among other non-Newtonian behaviors.
In \Fref{chap:lbm-stability}, the accuracy, stability, and efficiency of Lattice Boltzmann models in simulating non-Newtonian fluid flow was investigated.
Two different benchmark problems were chosen for the study: Poiseuille flow, and lid-driven square cavity flow, as Poiseuille flow is characterized by the development of velocity profiles and lid-driven cavity flow is characterized by the development of vortices.
Conclusions from the study were presented in \Fref{sec:numerical-study-conclusions}.
In the context of studying the accuracy, stability, and efficiency of Lattice Boltzmann models for simulating non-Newtonian flow, future work would include investigating the effect of the choice of time step size and simulation Mach number (as in~\cite{conrad2015accuracy}) on the convergence rate and accuracy of the results and LB methods presented in \Fref{sec:numerical-study}.
In addition, possible future work could include investigating other physical problems such as planar sudden expansion flow~\cite{tang2011binghami,chai2011multiple,chen2014simulations,wang2008lattice}, reentrant corner flow~\cite{gabbanelli2005lattice,yoshino2007numerical}, and flow around a cylinder~\cite{fallah2012multiple}.
In the context of simulating wellbore cementing, future work would include studying the accuracy, stability, and efficiency of Lattice Boltzmann models for simulating the thixotropic and time-dependent behavior of cement slurry~\cite{roussel2005steady}.

In \Fref{chap:lbm-cement}, the LBM algorithm for non-Newtonian flow is extended to simulate free surface flow.
The extended LBM model is then used to study wellbore cementing.
Future work for the wellbore cementing study would include developing a more physically realistic model.
Primary cementing can involve multiple phases and multiple components, namely the cement slurry, a spacer fluid, and drilling mud.
Possible future work would be to make the model more general so as to be capable of simulating multiphase multicomponent flow.
In addition, a more physically realistic model would simulate three-dimensional flow so as to be able to simulate eccentric annuli and three-dimensional rock formation imperfections.

\bibliographystyle{abbrvnat}
\bibliography{master}

\end{document}