### 2016-09-10 sigandnov.tex
wrote short significance and novelty paragraph as required for submission

\documentclass{article}

\begin{document}

\section*{Significance and novelty of this paper}

In this paper, it is shown that Lattice Boltzmann methods have certain advantages that can make them effective for simulating non-Newtonian flows, but that producing an accurate and stable solution is non-trivial depending on the physical problem and the non-Newtonian constitutive relationship of interest.
Past work has been done to access and improve both the accuracy and stability of Lattice Boltzmann methods, however, little has been done to compare their relative effectiveness for simulating non-Newtonian fluids in terms of accuracy, stability, and efficiency.
The significance of this work is that it aims to, through a numerical investigation, determine the strengths and weaknesses of various Lattice Boltzmann Methods when applied to simulating non-Newtonian fluid flow; and in addition, it aims to identify the mechanisms through which error and instability are introduced into the simulation so as to make recommendations about how new developments can be made for improving Lattice Boltzmann methods that are applied to non-Newtonian flow.
The novelty of this paper is considered to be: (1) a quantitative comparison of the computational cost for the BGK and MRT collision operators as applied to non-Newtonian flow, (2) the application of entropic filtering to non-Newtonian flow, and (3) an identification of some of the possible mechanisms that lead to error and instability in Lattice Boltzmann Methods applied to non-Newtonian flow that, to the author's knowledge, have not previously been discussed.

\end{document}