### 2017-03-02 response-to-reviewers.tex
copied template for cover letter to editor and response to reviewers. fixed grammatical and spelling errors in the text

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

\begin{document}

\section{Response to reviewers}

Thank you very much for your assistance in reviewing this paper. We are sure that with your help, this has become a much better work. We hope that the attached edited copy of "Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow" will answer any questions previously raised, and will be satisfactory for publication.

The following is a list of all comments recieved, and how they have been addressed in the updated version:

\end{document}