### 2017-03-03 cover-letter-to-editor.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

\begin{document}

Fernando Pinho
Universidade do Porto, Porto, Portugal

Dear Prof. Pinho:

Thank you very much for your assistance with our article titled "Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow" by Matthew Grasinger, Scott Overacker, and John C. Brigham for publication in the Journal of Non-Newtonian Fluid Mechanics. We are submitting to you the revised manuscript in which we have addressed the suggestions by the reviewers. In addition to the revised manuscript, we have included a document of our responses to the reviewers. Please let us know if any further modification or information is required.

All correspond should be addressed to:

John C. Brigham
blah blah blah

\end{document}