### 2017-03-03 cover-letter-to-editor.tex

\documentclass{letter}
\signature{John C. Brigham \\
	\vspace{0.1in}
	Associate Professor \\
	School of Engineering \\ \hspace{0.1in} and Computing Sciences}
\address{South Road\\
	Durham\\
	DH1 3LE \\ United Kingdom}
\begin{document}
	\begin{letter}{Fernando Pinho \\ Universidade do Porto, Porto, Portugal}
		\opening{Dear Prof. Pinho,}
		Thank you very much for your assistance with our article titled ``Numerical Investigation of the Accuracy, Stability, and Efficiency of Lattice Boltzmann Methods in Simulating Non-Newtonian Flow'' by Matthew Grasinger, Scott Overacker, and John C. Brigham for publication in the Journal of Non-Newtonian Fluid Mechanics. We are submitting to you the revised manuscript in which we have addressed the suggestions by the reviewers. In addition to the revised manuscript, we have included a document of our responses to the reviewers. Please let us know if any further modification or information is required.

		All correspond should be addressed to:

		John C. Brigham\\
		South Road\\
		Durham\\
		DH1 3LE\\
		john.brigham@durham.ac.uk

		\vspace{1in}
		\closing{Sincerely,}
	\end{letter}
\end{document}