### 2018-02-12 manuscript-srcs/src/cementing-notes.tex
submission accepted for publicaton

- turbulent flow is good for displacement
- typical pump rates include 2--15 (2, 5, 10, 15) bbl/min
-- 4-6 is most common
- typical yield stress 10 lbf / 100 sqft = approximately 5 pascals
-- range from 2.5 to 30 pascals
- typical plastic viscosities are 10-100 cP
-- 1 Pa s = 10 P = 1000 cP

- standard casing sizes 4.5" -- 20" in outside diameter
- corresponding annuli areas: 12 -- 48 sqin 
                              0.08 -- 0.333 sqft
- 1 bbl = 5.6145833 cft
- 2 bbl/min   = 0.18715 cft / sec
  15 bbl/min  = 1.40365 cft / sec

- velocities range: 0.562 -- 17.5 ft/sec
- dynamic visco range: 0.000208854 -- 0.00208854 lbf s / ft^2
- slurry density range: 12.0 -- 18.0 lb/gal
                        90   -- 135 lb/cft
- kinematic visco range: 1.5e-6 -- 2.32e-5 ft s
- Re range: